Eviction and Crime: Evidence from Boston

### 2023-04-12 writing/paper/main.tex
Updates from Overleaf

\begin{titlepage}
\title{Eviction and Crime: Evidence from Boston}
\author{Arjun Shanmugam}
\date{\today}
\maketitle
\begin{abstract}
\noindent This paper investigates the effects of eviction on the frequency of crime incident responses in the immediate vicinity of a property. I use a difference-in-differences research design based on the staggered conclusion of eviction cases in Boston, Massachusetts between June 2019 and March 2020. Using properties where eviction was unsuccessful as the counterfactual for properties where eviction was successful, I find negative effects of eviction on crime incident responses which persist for about eleven months after case resolution. The magnitudes of these effects are approximately 0.5 percent of the mean number of crimes per month in the immediate vicinity of a property in 2017. Placebo tests using larceny and burglary incident responses as an outcome variable suggest that these effects are driven by the movement of individuals involved in crime incidents outside the radii of properties where evictions are successful. Analysis of treatment effects using only cases resolved before the beginning of the pandemic provides evidence against the hypothesis that my results are driven by a pandemic-related shift in characteristics of evicted tenants.

@@

This paper seeks to reinforce our knowledge of eviction's effects on small communities by estimating the effect of eviction on the frequency of crime incidents within 500 meters of a property. Empirical research of eviction and its impacts faces two main roadblocks \citep{collinson_eviction_2022}. The first is the difficulty of conducting analysis at the individual- or property-level. Eviction case records are often scattered across disjoint public and private organizations and difficult to obtain in bulk. It is also challenging to link eviction case records to individual- and property-level outcomes. I overcome this barrier by obtaining crime incident-level data from the Boston Police Department and eviction case-level data from MassLandlords, a trade association of landlords in Massachusetts. Armed with records of every eviction case resolved in Massachusetts Housing Court since June 2019 and every Boston Police Department crime incident response since August 2015, I spatially join each eviction record from Boston with any crime incident responses which occurred within 500 meters of the associated property. Using this comprehensive crime incident response data, I produce a panel that allows me to observe crime incident counts near each property for several years before and after case resolution.

@@

ADD DISCUSISON NOF PANDEMIC: ONLY THE MOST SERIOUS EVICTIONS PROCEEDED, BIASING EVICTEES TO BE MORE SERIOUS OFFENDERS

@@

        \subsubsection{The Eviction Process}
        Tenants in Massachusetts can rent from landlords under the terms of a lease or without one. If a tenant is renting from a landlord without a lease, their tenancy is \textit{at will}. The nature of the agreement between the tenant and the landlord—the \textit{landlord-tenant agreement}—determines the steps that the landlord must take to evict the tenant.

@@

    \subsection{Crime Incident Responses}

@@

    \subsection{Evictions Data}

        \newpage
        \begin{figure}[H]
            \centering
            \includegraphics{output/summary_statistics/figures/evictions_map.png}
            \caption{Spatial Incidence of Eviction}
            \label{fig:my_label}
        \end{figure}

        \begin{landscape}
        \begin{figure}[H]
            \centering
            \includegraphics[scale=1.25]{output/summary_statistics/figures/filings_over_time.png}
            \caption{Eviction Filings Over Time in Boston}
            \label{fig:my_label}
        \end{figure}
\end{landscape}
        \begin{table}[H]
            \centering
            \small
            \input{output/summary_statistics/tables/treatment_timings.tex}
            \caption{Distribution of Eviction Filings and Outcomes}
            \label{tab:my_label}
        \end{table}

    \subsection{Summary Statistics}
         \begin{table}[H]
            \centering
            \small
            \input{output/summary_statistics/tables/summary_statistics}
            \caption{Summary Statistics}
            \label{tab:table_1}
        \end{table}
        \newpage

>>>>

\begin{titlepage}
\title{Eviction and Crime: Evidence from Boston}
\author{Arjun Shanmugam}
\date{\today}
\maketitle
\begin{abstract}
\noindent This paper investigates the effects of eviction on the frequency of crime incident responses in the immediate vicinity of a property. I use a doubly robust difference-in-differences research design based on the staggered conclusion of eviction cases in Boston, Massachusetts between April 2019 and March 2020. Using properties where eviction was unsuccessful as the counterfactual for properties where eviction was successful, I find negative effects of eviction on crime incident responses which persist for 3 years after case resolution. Point estimates of the magnitudes of these effects are approximately between 2 and 4 percent of the mean number of crimes per month in the immediate vicinity of a property in 2017. Effects are driven by impacts during warm months.

Placebo tests using larceny and burglary incident responses as an outcome variable suggest that these effects are driven by the movement of individuals involved in crime incidents outside the radii of properties where evictions are successful. Analysis of treatment effects using only cases resolved before the beginning of the pandemic provides evidence against the hypothesis that my results are driven by a pandemic-related shift in characteristics of evicted tenants.

@@

This paper seeks to reinforce our knowledge of eviction's effects on small communities by estimating the effect of eviction on the frequency of crime incidents within 500 meters of a property. Empirical research of eviction and its impacts faces two main roadblocks \citep{collinson_eviction_2022}. The first is the difficulty of conducting analysis at the individual- or property-level. Eviction case records are often scattered across disjoint public and private organizations and difficult to obtain in bulk. It is also challenging to link eviction case records to individual- and property-level outcomes. I overcome this barrier by obtaining crime incident-level data from the Boston Police Department (BPD) and eviction case-level data from MassLandlords, a trade association of landlords in Massachusetts. Armed with records of every eviction case resolved in Massachusetts Housing Court since June 2019 and every BPD crime incident response since August 2015, I spatially join each eviction record from Boston with any crime incident responses which occurred within 500 meters of the associated property. Using this comprehensive crime incident response data, I produce a panel that allows me to observe crime incident counts near each property for several years before and after case resolution.

@@

@@

        \subsubsection{The Eviction Process in Massachusetts}
        Tenants in Massachusetts can rent from landlords under the terms of a lease or without one. If a tenant is renting from a landlord without a lease, their tenancy is \textit{at will}. The nature of the agreement between the tenant and the landlord—the \textit{landlord-tenant agreement}—determines the steps that the landlord must take to evict the tenant.

@@

    \subsection{Crime in Boston}

    1. How much crime is in Boston compared to most cities?
    2. Where is crime worst in Boston?
    3. When does the boston police department respond to an incident?

@@

    \subsection{Evictions Data}
    I obtain records of every summary process case filed in Boston since April 2019 from MassLandlords, a trade association of landlords in Massachusetts. At the beginning of the pandemic, MassLandlords developed a sophisticated system for manually collecting and programatically scraping court dockets. Crucially, each record includes the resolution of the case, the last date on the case docket, and the address of the disputed property. Each record also includes details such as the court in which the case was initially filed, the duration of the case, whether the tenant and landlord had attorneys, and the type of notice to quit that was initially filed by the landlord. I use a paid geocoding service known as Geocodio to obtain highly accurate latitude and longitude coordinates for each property.

    I restrict my sample in 4 ways. First, I drop all records which could not be geocoded. Second, I drop the less than one percent of records for which a latest docket date could not be found. Third, I drop all cases which were resolved through mediation. Lastly, I drop cases for which a judgment could not be found. Out of \textcolor{red}{X} cases filed in Boston between April 2019 and March 2020, I am left with a sample of \textcolor{red}{Y} cases. Figure 1 displays the number of evictions in my sample filed during each month.

        \begin{figure}[H]
            \centering
            \includegraphics{output/summary_statistics/figures/filings_over_time.png}
            \caption{Eviction Filings Over Time in Boston}
            \label{fig:my_label}
        \end{figure}

    \subsection{Crime Incidents Data}
    I obtain incident-level records of every crime incident to which BPD offers responded to from August 2015 to present. Each record includes the a date, latitude and longitude coordinates, and an offense code indicating the type of crime incident that occurred. I drop all records missing any of these three variables.

    \subsection{Census Tract Characteristics Data}
    I obtain time-invariant census tract level characteristics of the properties in my sample from Opportunity Insights \citep{chetty_impacts_2018}. This data includes information such as population density, job density, median household income, and poverty rate. Figure 2 plots the locations of the evictions in my sample, coloring Boston's census tracts according to their poverty rates. It shows that the density of eviction filings tends to be higher in poor areas.

            \begin{figure}[H]
            \centering
            \includegraphics{output/summary_statistics/figures/evictions_map.png}
            \caption{Spatial Incidence of Eviction}
            \label{fig:my_label}
        \end{figure}

    \subsection{Merged Dataset}
    My final dataset is a panel dataset of nearby crime incident counts at the property month-level. To produce this dataset, I use GIS tools to draw circles of radius 500 meters around each property in my dataset of evictions. I match each property with all crimes that fall within its radius. Then, for each property, I count the number of crimes that occurred within its radius during each month from August 2015 to present.

    Table 1 contains descriptive statistics for the panel dataset.

         \begin{table}[H]
            \centering
            \small
            \input{output/summary_statistics/tables/summary_statistics}
            \caption{Summary Statistics}
            \label{tab:table_1}
        \end{table}

       \begin{table}[H]
            \centering
            \small
            \input{output/summary_statistics/tables/treatment_timings.tex}
            \caption{Distribution of Eviction Filings and Outcomes}
            \label{tab:my_label}
        \end{table}
